Follow-up Samples

### 2016-12-07 submission/manuscript_outline_20161024.Rmd
update introduction text

Colorectal cancer (CRC) continues to be a leading cause of cancer related deaths and is the second most common cancer death among men aged 40-79 years of age (**insert citation**).  Over the last few years death due to the disease has seen a significant decrease thanks mainly to improvements in screening (**insert citation**).  However, despite this giant improvement there are still approximately 50,000 deaths from the disease a year (**insert citation**).  It is estimated that around 5-10% of all CRCs can be explained by autosomal dominant inheritance (**insert citation**). The vast majority of CRCs are not inherited and the exact etiology to disease is not well worked out (**insert citation**).

Recent reports have found that the microbiome between those with and without CRC are different and that bacteia associated with CRC tissue is different that that found on the normal mucosa (**Insert citations**).  Further, there is evidence to suggest that more inflammatory mouth-associated bacteria are more prevalent in the guts of those with CRC (**insert citation**).  This has led to the hypothesis that these bacteria replace traditional residents that produce metabolites such as butyrate (**insert citation**) and lead to a more inflammatory state (**insert citation**).  This inflammatory state allows for more of these types of bacteria to populate the gut and leads to neoplasia and eventually CRC (**insert citation**).

Building upon these findings a number of groups, including our own, have investigated using the bacterial microbiome as a potential biomarker for adenoma and CRC detection (**insert citations**).  These collectively show that a reasonable area under the curve (AUC) can be obtained when using the stool bacterial microbiome to classify disease state (**insert citation**).  Further, a number of groups have extended this observation to show that using either specific bacterial species or the bacterial microbiome as a whole paired with the Fecal Immunochemical Test (FIT) can either increase the AUC of the model (**insert citations**).

In this study we further refine the random forest model by including more background information on the indivdiuals within it. We use a rigourous and accepted approach for machine learning training and testing to validate this model (**insert citation**).  We then apply it to individuals in which we have samples before and after treatment to assess how well the model adjusts their probability of having an adenoma or carcinoma.  Finally, we investigate which specific Operational Taxonomic Units (OTUs) are most affected by treatment.  Within this study lesion refers to both adenoma and carcinoma.

>>>>

Colorectal cancer (CRC) continues to be a leading cause of cancer related deaths and is the second most common cancer death among men aged 40-79 years of age (**insert citation**).  Over the last few years death due to the disease has seen a significant decrease thanks mainly to improvements in screening (**insert citation**).  However, despite this giant improvement there are still approximately 50,000 deaths from the disease a year (**insert citation**).  It is estimated that around 5-10% of all CRCs can be explained by autosomal dominant inheritance (**insert citation**). The vast majority of CRCs are not inherited and the exact etiology to disease has not been well worked out (**insert citation**).

Recent reports have found that the microbiome between those with and without CRC are different and that bacteia associated with CRC tissue is different than that found on the normal mucosa (**Insert citations**).  Further, there is evidence to suggest that more inflammatory mouth-associated bacteria are more prevalent in the guts of those with CRC (**insert citation**).  This has led to the hypothesis that these bacteria replace traditional residents that produce metabolites such as butyrate (**insert citation**) and lead to a more inflammatory state (**insert citation**).  This inflammatory state allows for more of these types of bacteria to populate the gut and leads to neoplasia and eventually CRC (**insert citation**).

Building upon these findings a number of groups, including our own, have investigated using the bacterial microbiome as a potential biomarker for adenoma and CRC detection (**insert citations**).  These collectively show that a reasonable area under the curve (AUC) can be obtained when using the stool bacterial microbiome to classify disease state (**insert citation**).  Further, a number of groups have extended this observation to show that using either specific bacterial species or the bacterial microbiome as a whole paired with the Fecal Immunochemical Test (FIT) can increase the AUC of the model (**insert citations**).

In this study we further refine the random forest model by including more background information on the individuals within it. We used a rigourous and accepted approach for machine learning training and testing to validate this model (**insert citation**).  We then applied it to individuals in which we have samples before and after treatment to assess how well the model adjusts their probability of having an adenoma or carcinoma.  Finally, we investigated which specific Operational Taxonomic Units (OTUs) are most affected by treatment.  Within this study lesion refers to both adenoma and carcinoma.

### 2016-12-07 submission/manuscript_outline_20161024.tex
update introduction text

Colorectal cancer (CRC) continues to be a leading cause of cancer
related deaths and is the second most common cancer death among men aged
40-79 years of age (\textbf{insert citation}). Over the last few years
death due to the disease has seen a significant decrease thanks mainly
to improvements in screening (\textbf{insert citation}). However,
despite this giant improvement there are still approximately 50,000
deaths from the disease a year (\textbf{insert citation}). It is
estimated that around 5-10\% of all CRCs can be explained by autosomal
dominant inheritance (\textbf{insert citation}). The vast majority of
CRCs are not inherited and the exact etiology to disease is not well
worked out (\textbf{insert citation}).

Recent reports have found that the microbiome between those with and
without CRC are different and that bacteia associated with CRC tissue is
different that that found on the normal mucosa (\textbf{Insert
citations}). Further, there is evidence to suggest that more
inflammatory mouth-associated bacteria are more prevalent in the guts of
those with CRC (\textbf{insert citation}). This has led to the
hypothesis that these bacteria replace traditional residents that
produce metabolites such as butyrate (\textbf{insert citation}) and lead
to a more inflammatory state (\textbf{insert citation}). This
inflammatory state allows for more of these types of bacteria to
populate the gut and leads to neoplasia and eventually CRC
(\textbf{insert citation}).

Building upon these findings a number of groups, including our own, have
investigated using the bacterial microbiome as a potential biomarker for
adenoma and CRC detection (\textbf{insert citations}). These
collectively show that a reasonable area under the curve (AUC) can be
obtained when using the stool bacterial microbiome to classify disease
state (\textbf{insert citation}). Further, a number of groups have
extended this observation to show that using either specific bacterial
species or the bacterial microbiome as a whole paired with the Fecal
Immunochemical Test (FIT) can either increase the AUC of the model
(\textbf{insert citations}).

In this study we further refine the random forest model by including
more background information on the indivdiuals within it. We use a
rigourous and accepted approach for machine learning training and
testing to validate this model (\textbf{insert citation}). We then apply
it to individuals in which we have samples before and after treatment to
assess how well the model adjusts their probability of having an adenoma
or carcinoma. Finally, we investigate which specific Operational
Taxonomic Units (OTUs) are most affected by treatment. Within this study
lesion refers to both adenoma and carcinoma.

>>>>

Colorectal cancer (CRC) continues to be a leading cause of cancer
related deaths and is the second most common cancer death among men aged
40-79 years of age (\textbf{insert citation}). Over the last few years
death due to the disease has seen a significant decrease thanks mainly
to improvements in screening (\textbf{insert citation}). However,
despite this giant improvement there are still approximately 50,000
deaths from the disease a year (\textbf{insert citation}). It is
estimated that around 5-10\% of all CRCs can be explained by autosomal
dominant inheritance (\textbf{insert citation}). The vast majority of
CRCs are not inherited and the exact etiology to disease has not been
well worked out (\textbf{insert citation}).

Recent reports have found that the microbiome between those with and
without CRC are different and that bacteia associated with CRC tissue is
different than that found on the normal mucosa (\textbf{Insert
citations}). Further, there is evidence to suggest that more
inflammatory mouth-associated bacteria are more prevalent in the guts of
those with CRC (\textbf{insert citation}). This has led to the
hypothesis that these bacteria replace traditional residents that
produce metabolites such as butyrate (\textbf{insert citation}) and lead
to a more inflammatory state (\textbf{insert citation}). This
inflammatory state allows for more of these types of bacteria to
populate the gut and leads to neoplasia and eventually CRC
(\textbf{insert citation}).

Building upon these findings a number of groups, including our own, have
investigated using the bacterial microbiome as a potential biomarker for
adenoma and CRC detection (\textbf{insert citations}). These
collectively show that a reasonable area under the curve (AUC) can be
obtained when using the stool bacterial microbiome to classify disease
state (\textbf{insert citation}). Further, a number of groups have
extended this observation to show that using either specific bacterial
species or the bacterial microbiome as a whole paired with the Fecal
Immunochemical Test (FIT) can increase the AUC of the model
(\textbf{insert citations}).

In this study we further refine the random forest model by including
more background information on the individuals within it. We used a
rigourous and accepted approach for machine learning training and
testing to validate this model (\textbf{insert citation}). We then
applied it to individuals in which we have samples before and after
treatment to assess how well the model adjusts their probability of
having an adenoma or carcinoma. Finally, we investigated which specific
Operational Taxonomic Units (OTUs) are most affected by treatment.
Within this study lesion refers to both adenoma and carcinoma.
